\documentclass[a4paper, 12pt]{article}

\usepackage{upquote}

\begin{document}

\title{Android Notes}
\author{Nazmus Saquib}

\maketitle
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Project Files}
\begin{description}
	\item[src] this folder contains all of our java files
	\item[gen] this folder cotains generated files, we should not need to modify these
	\item[res] folder deals with images, resoources etc. the drawable folders can contain images of different resolution; SDK will use the most appropriate one
	\item[res/layout] contains layout specification of app
	\item[res/values] a number of values folders can be seen, these can be later referenced
	\item[AndroidManifest.xml] maintains everything, boss mode
	\item[res/layout-land] custom folder, layout in case our phone is tilted
	\item[res/raw] custom folder, contains audio
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source File Syntax}
\begin{itemize}
	\item \verb|setContentView(R.layout.activity_main)| : sets the layout according to res/layout/activity\_main.xml
\end{itemize}

%-----------------------------------------------%
\subsection{Connection between XML and Java}
\emph{bAdd} is the id as defined in the XML file.
\begin{verbatim}
Button bAdd;
bAdd = (Button) findViewById(R.id.bAdd);
\end{verbatim}

%-----------------------------------------------%
\subsection{Button Click}
Note that argument of \emph{setText} can not be only number.
\begin{verbatim}
add.setOnClickListener(new View.onClickListener(){
	public void onClick(View v){
		counter++;
		dsiplay.setText("Your Total is: " + counter);
	}
});
\end{verbatim}

%-----------------------------------------------%
\subsection{Action Listener Example}
\begin{verbatim}
public void sendMessage(View view){

}
\end{verbatim}
Note that in order to match with the \emph{android:onClick} attribute, the signature must have the following properties:
\begin{itemize}
	\item public method
	\item return value of void
	\item have only one parameter, namely of \emph{View} type
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Layout File for XML}
\begin{itemize}
	\item if a tag does not have another tags inside of it, we can simply use short form of tag close
	\item \verb|android:layout_width="fill_parent"|	: fills the size of the parent
	\item \verb|android:layout_width="wrap_content"| : uses just what is needed
	\item \verb|android:text="@string/ref"| : res/values/strings.xml contains the string ref, defined as say \verb|<string name="ref">this is a reference</string>|, so ref has the value within tags
	\item \verb|android:text="string"| : puts the string within quotes directly
	\item \verb|android:textSize="40sp"| : text size, dp or px should not be used, basically we should use sp for fonts and dp for everything else
	\item \verb|android:layout_gravity="center"| put component at the center, notice we'll need wrap\_content. Otherwise we can also specify \emph{gravity} to be center
	\item \verb|android:@+id/tvDisplay| we are setting up and \emph{id}, and that is \emph{tvDisplay}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{UI Basics}
\begin{itemize}
	\item \emph{ViewGroup}s are container of views such as \emph{LinearLayout}s, \emph{RelativeLayout}s.
	\item \emph{View}s are objects such as \emph{Button}s, \emph{TextView}s etc.
\end{itemize}

%-----------------------------------------------%
\subsection{ViewGroup}
\begin{description}
	\item[LinearLayout] components appear in the order they are written in the XML file.
\end{description}

%-----------------------------------------------%
\subsection{Controls}
\begin{description}
	\item[EditText] user editable text.	
\end{description}

%-----------------------------------------------%
\subsection{Control Attributes}
\begin{description}
	\item[android:id] unique identifier for the view. The \emph{@} sign is required whenever we are referring to any resource object. Plus signs are necessary when creating a resource ID for the first time.
	\item[android:hint] default text to show when the control is blank.
	\item[android:layout\_weight] indicates how much of a free space each component should consume. Once we give a weight, it is better to set \emph{layout\_width} to 0dp, as putting anything else would be irrelevant because the width has to be calculated to match the weight.
	\item[android:onClick] the value is the name of a function in the activity that gets called when the control is clicked.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Activity}
Can be simply thought of as a screen. All subclasses of activity must implement \emph{onCreate} method. All activities must be declared in our manifest file using \verb|<activity>| tag. Every activity is invoked by an intent, we can get the intent by invoking \emph{getIntent} method.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intent}
An \emph{Intent} is an object that provides runtime binding between seperate components, such as two activities. We need to import the intent class - \verb|import android.component.Intent|. An Intent can carry a collection of various data types as key-value pairs called \emph{extras}. The \emph{putExtra} method takes the key as first parameter and the value as the second parameter. An activity is started by invoking the \emph{startActivity} method with an Intent as parameter.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Command Line}
%-----------------------------------------------%
\subsection{Genymotion}
Make sure to use the path to genymotion, better yet, add to profile / bashrc.
\begin{itemize}
	\item \verb|VBoxManage list vms| shows device name and id
	\item \verb|player --vm-name name-of-device| launches virtual device
	\item \verb|adb install path-to-apk-in-bin| installs apk
	\item \verb|adb uninstall packagename| uninstalls apk
\end{itemize}
%-----------------------------------------------%
\subsection{SDK}
All the tools can be found in \emph{sdk/tools}, the adb has been moved.
\begin{itemize}
	\item \verb|./android avd| launches avd manager, note that once we are using genymotion we won't need this
	\item \verb|./emulator -avd emulator-name| launches the emulator
	\item \verb|adb -s emulator-5554 install apk-path-name| installs the apk in avd, \emph{5554} is the emulator number
	\item \verb|./android list targets| lists virtual devices
	\item The following creates a project with supplied args:
		\begin{verbatim}
			./android create project \
			--target <target_ID> \
			--name <your_project_name> \
			--path path/to/your/project \
			--activity <your_activity_name> \
			--package <your_package_namespace>
		\end{verbatim}
	\item \verb|ant debug| from the project root folder builds the project with debugging
\end{itemize}
%-----------------------------------------------%
\subsection{ADB push}
We can transfer files from pc to device by:
\begin{verbatim}
adb push path/to/file/pc /storage/sdcard0/downloads
\end{verbatim}
Here we have assumed we want to put files in the downloads folder.
One way to check the path of android device is to use bluetooth history.
If we want to transfer all files of a directory, we can use ``.'' in place of pc path.

\end{document}
